---
title: "BMI Dataset"
author: "Amanda Carrico"
date: "2024-05-29"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    number_sections: no
    theme: cerulean
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  word_document:
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```

  

# Load in data (put together any dataset with BMI that had over 100-200 patients)

## Bladder Urothelial Carcinoma TCGA/Firehose Legacy
```{r}
Bladder_data_clinical <- read_csv("Bladder Urothelial_data_clinical_patient.csv")
bladder_data_patient <- Bladder_data_clinical %>% select(PATIENT_ID, WEIGHT, HEIGHT, SEX, RACE, ETHNICITY, SITE_OF_TUMOR_TISSUE, METASTATIC_SITE_PATIENT, TOBACCO_SMOKING_HISTORY_INDICATOR, SMOKING_PACK_YEARS, AGE, OS_STATUS, OS_MONTHS) %>% mutate(BMI = as.numeric(WEIGHT)/(as.numeric(HEIGHT)/100)^2, OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), METASTATIC_SITE_PATIENT = ifelse(METASTATIC_SITE_PATIENT == "[Not Available]" | METASTATIC_SITE_PATIENT == "None", NA, METASTATIC_SITE_PATIENT), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN", "Asian", RACE)), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ETHNICITY), SMOKING_HISTORY = ifelse(TOBACCO_SMOKING_HISTORY_INDICATOR == "[Not Available]", NA, TOBACCO_SMOKING_HISTORY_INDICATOR), SMOKING_PACK_YEARS = ifelse(SMOKING_PACK_YEARS == "[Not Available]", NA, SMOKING_PACK_YEARS)) %>% select(-c(WEIGHT, HEIGHT, TOBACCO_SMOKING_HISTORY_INDICATOR)) %>% filter(!is.na(BMI)) %>% mutate(BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))))


Bladder_data_sample <- read_csv("Bladder Urothelial_data_clinical_sample.csv")
Bladder_data_sample <- Bladder_data_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n(), TMB = TMB_NONSYNONYMOUS, Study = "Bladder Urothelial Carcinoma TCGA/Firehose Legacy", TUMOR_PURITY = NA, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-TMB_NONSYNONYMOUS)
Bladder_data <- merge(bladder_data_patient, Bladder_data_sample, by = c("PATIENT_ID")) %>% filter(!(PATIENT_ID == "TCGA-DK-A1A6" & TMB == NA))

head(Bladder_data) 
```

## Glioblastoma CPTAC/Cell 2021
```{r}
Glioblastoma_data_clinical <- read_csv("Glioblastoma_data_clinical_patient.csv")
Glioblastoma_data_clinical <- Glioblastoma_data_clinical %>% select(PATIENT_ID, BMI, AGE, SEX, RACE, ETHNICITY, VITAL_STATUS, FOLLOW_UP_PERIOD, SMOKING_HISTORY) %>% mutate(OS_STATUS = ifelse(VITAL_STATUS == "Deceased", 1, 0), OS_MONTHS = substr(FOLLOW_UP_PERIOD, 1,2), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "White", "WHITE", RACE)), SMOKING_PACK_YEARS = NA, SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ifelse(ETHNICITY == "Hispanic or Latino", "HISPANIC OR LATINO", ifelse(ETHNICITY == "Not-Hispanic or Latino", "NOT HISPANIC OR LATINO", ETHNICITY))), BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese")))) %>% select(-c(FOLLOW_UP_PERIOD, VITAL_STATUS))


Glioblastoma_data_sample <- read_csv("Glioblastoma_data_clinical_sample.csv")
Glioblastoma_data_sample <- Glioblastoma_data_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS, TUMOR_SITE_CURATED) %>% mutate(TMB = TMB_NONSYNONYMOUS, SITE_OF_TUMOR_TISSUE = TUMOR_SITE_CURATED, Study = "Glioblastoma CPTAC/Cell 2021", METASTATIC_SITE_PATIENT = NA, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-c(TUMOR_SITE_CURATED, TMB_NONSYNONYMOUS)) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n(), TUMOR_PURITY = NA) # checked and no patients with multiple samples

Glioblastoma_data <- merge(Glioblastoma_data_clinical, Glioblastoma_data_sample, by = c("PATIENT_ID"))

head(Glioblastoma_data)
```

## Colorectal Cancer MSK/JNCI 2021
```{r}
Colorectal_data_clinical <- read_csv("Colorectal_data_clinical_patient.csv")
Colorectal_data_clinical <- Colorectal_data_clinical %>% select(PATIENT_ID, BMI, RACE, SEX, AGE_AT_DX, OS_MET_STATUS, OS_MET_MONTHS, SMOKING_HISTORY) %>% mutate(OS_STATUS = ifelse(OS_MET_STATUS == "1:DECEASED", 1, 0), OS_MONTHS = OS_MET_MONTHS, AGE = AGE_AT_DX, ETHNICITY = NA, RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN-FAR EAST/INDIAN SUBCONT", "Asian", ifelse(RACE == "UNKNOWN_OTHER", "Other", ifelse(RACE == "NATIVE AMERICAN-AM IND/ALASKA", "AMERICAN INDIAN OR ALASKA NATIVE", RACE)))), SEX = ifelse(SEX == "[Not Available]", NA, SEX), SMOKING_PACK_YEARS = NA, BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30 , "Overweight", "Obese")))) %>% select(-c(AGE_AT_DX, OS_MET_MONTHS, OS_MET_STATUS)) # met status and months also contained - survival from diagnosis of METS (but not the primary)

Colorectal_data_sample <- read_csv("Colorectal_data_clinical_sample.csv")
#Colorectal_data_sample %>% select(SAMPLE_ID, PATIENT_ID, PRIMARY_TUMOR_LOCATION, CANCER_TYPE, METASTATIC_SITE, CVR_TMB_SCORE, TUMOR_PURITY) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n())  #checked and no multiple patients

Colorectal_data_sample <- Colorectal_data_sample %>% select(SAMPLE_ID, PATIENT_ID, PRIMARY_TUMOR_LOCATION, METASTATIC_SITE, CANCER_TYPE, CVR_TMB_SCORE, TUMOR_PURITY) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n(), TMB = CVR_TMB_SCORE, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE), SITE_OF_TUMOR_TISSUE = PRIMARY_TUMOR_LOCATION, Study = "Colorectal Cancer MSK/JNCI 2021", METASTATIC_SITE_PATIENT = ifelse(METASTATIC_SITE == "Not Applicable", NA, METASTATIC_SITE)) %>% select(-c(PRIMARY_TUMOR_LOCATION, CVR_TMB_SCORE, METASTATIC_SITE))

Colorectal_data <- merge(Colorectal_data_clinical, Colorectal_data_sample, by = c("PATIENT_ID"))

head(Colorectal_data)
```

## Esophagogastric Cancer MSK/J Natl Cancer Inst 20
```{r}
Esophagogastric_data_clinical <- read_csv("Esophagogastric_data_clinical_patient.csv")

Esophagogastric_data_clinical <- Esophagogastric_data_clinical %>% select(PATIENT_ID, AGE_AT_DIAGNOSIS, OS_STATUS, OS_MONTHS, BMI_CATEGORIES, SEX, RACE, ETHNICITY) %>% mutate(BMI = NA, BMI_CATEGORIES = ifelse(BMI_CATEGORIES == "Unknown", NA, BMI_CATEGORIES), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "White", "WHITE", ifelse(RACE == "Black or African AMERICAN", "BLACK OR AFRICAN AMERICAN", RACE))), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ifelse(ETHNICITY == "Not Hispanic or Latino", "NOT HISPANIC OR LATINO", ifelse(ETHNICITY == "Hispanic or Latino", "HISPANIC OR LATINO", ifelse(ETHNICITY == "not reported", "Unknown", ETHNICITY)))), SMOKING_HISTORY = NA, SMOKING_PACK_YEARS = NA, OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), AGE = AGE_AT_DIAGNOSIS) %>% select(-AGE_AT_DIAGNOSIS) # could not find cutoffs for BMI categories

Esophagogastric_data_sample <- read_csv("Esophagogastric_data_clinical_sample.csv")

Esophagogastric_data_sample <- Esophagogastric_data_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS, PRIMARY_SITE, METASTATIC_SITE, TUMOR_PURITY, TMB_NONSYNONYMOUS) %>% mutate(SITE_OF_TUMOR_TISSUE = PRIMARY_SITE, TMB = TMB_NONSYNONYMOUS, METASTATIC_SITE_PATIENT = ifelse(METASTATIC_SITE == "Not Applicable", NA, METASTATIC_SITE)) %>% select(-c(PRIMARY_SITE, TMB_NONSYNONYMOUS, METASTATIC_SITE)) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n(), Study = "Esophagogastric Cancer MSK/J Natl Cancer Inst 2023") 

# checked and no multiple samples - filter Code_frequency > 1

Esophagogastric_data <- merge(Esophagogastric_data_clinical, Esophagogastric_data_sample, by = c("PATIENT_ID"))

head(Esophagogastric_data)
```

## Hepatocellular Carcinoma (MSK 2024)
```{r}
Hepatocellular <- read_csv("Hepatocellular Carcinoma.csv") # no option to download full file
# `Fraction Genome Altered` can be added after figure out what to do about others
Hepatocellular_data <- Hepatocellular %>% select(`Patient ID`, `Sample ID`, `Cancer Type`, BMI, Age, Race, Sex, Ethnicity, `Metastatic Site`, `Overall Survival (Months)`, `Overall Survival Status`, `Primary Tumor Site`, `TMB (nonsynonymous)`, `Tumor Purity`) %>% mutate(PATIENT_ID = `Patient ID`, SAMPLE_ID = `Sample ID`, SEX = Sex, RACE = ifelse(Race == "[Not Available]", NA, ifelse(Race == "White", "WHITE", ifelse(Race == "Black", "BLACK OR AFRICAN AMERICAN", Race))), ETHNICITY = ifelse(Ethnicity == "Non-Spanish; Non-Hispanic" | Ethnicity == "South/Central America (except Brazil)" | Ethnicity == "Dominican Republic" | Ethnicity == "Cuban" | Ethnicity == "Other Spanish/Hispanic(incl European; excl Dom Rep" | Ethnicity == "Puerto Rican", "NOT HISPANIC OR LATINO", ifelse(Ethnicity == "Spanish  NOS; Hispanic NOS, Latino NOS", "HISPANIC OR LATINO", ifelse(Ethnicity == "Unknown whether Spanish or not", "Unknown", Ethnicity))), BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), OS_STATUS = ifelse(`Overall Survival Status` == "1:DECEASED", 1, 0), OS_MONTHS = `Overall Survival (Months)`, SMOKING_HISTORY = NA, SMOKING_PACK_YEARS = NA, CANCER_TYPE = `Cancer Type`, SITE_OF_TUMOR_TISSUE = `Primary Tumor Site`, AGE = Age, TMB = `TMB (nonsynonymous)`, TUMOR_PURITY = `Tumor Purity`, TMB = `TMB (nonsynonymous)`, METASTATIC_SITE_PATIENT = ifelse(`Metastatic Site` == "Not Applicable", NA, `Metastatic Site`), Study = "Hepatocellular Carcinoma (MSK 2024)")  %>% select(-c(`Patient ID`, `Sample ID`, `Cancer Type`, Age, `Metastatic Site`, `Overall Survival (Months)`, `Overall Survival Status`, Sex, Race, Ethnicity, `Primary Tumor Site`, `TMB (nonsynonymous)`, `Tumor Purity`)) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n()) 

head(Hepatocellular_data)
```

Problem: only see fraction genome altered when download the set from select on site (not seen when combine the two files in the official folder - I thought perhaps it was under a different variable name but checked and not found)


## Cervical Squamous Cell Carcinoma and Endocervical Adenocarcinoma (TCGA, Firehose Legacy)

```{r}
cervical_clinical_patient <- read_csv("cervical_data_clinical_patient.csv")

cervical_clinical_patient <- cervical_clinical_patient %>% select(PATIENT_ID, SEX, HEIGHT, WEIGHT, RACE, ETHNICITY, AGE, SITE_OF_TUMOR_TISSUE, OS_STATUS, OS_MONTHS, TOBACCO_SMOKING_HISTORY_INDICATOR, SMOKING_PACK_YEARS) %>% mutate(BMI = as.numeric(WEIGHT)/(as.numeric(HEIGHT)/100)^2, BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), METASTATIC_SITE_PATIENT = NA, SMOKING_PACK_YEARS = ifelse(SMOKING_PACK_YEARS == "[Not Available]", NA, SMOKING_PACK_YEARS), SMOKING_HISTORY = ifelse(TOBACCO_SMOKING_HISTORY_INDICATOR == "[Not Available]", NA, TOBACCO_SMOKING_HISTORY_INDICATOR), TUMOR_PURITY = NA, RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN", "Asian", RACE)), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ETHNICITY), Study = "Cervical Squamous Cell Carcinoma and Endocervical Adenocarcinoma (TCGA, Firehose Legacy)", OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0)) %>% select(-c(HEIGHT, WEIGHT, TOBACCO_SMOKING_HISTORY_INDICATOR))

cervical_clinical_sample <- read_csv("cervical_data_clinical_sample.csv")

cervical_clinical_sample <- cervical_clinical_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(TMB = TMB_NONSYNONYMOUS, Code_frequency= n(), CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-TMB_NONSYNONYMOUS)

cervical_data <- merge(cervical_clinical_patient, cervical_clinical_sample, by = "PATIENT_ID") %>% filter(!(PATIENT_ID == "TCGA-HM-A6W2" & CANCER_TYPE == "Cervical Cancer, NOS" | PATIENT_ID == "TCGA-UC-A7PG" & CANCER_TYPE == "Cervical Cancer, NOS"))
head(cervical_data)
# same patient, different cancer type - two cases of this where one is a more specific form of cancer perhaps diagnosed later but all other numbers are the same
```

## Intrahepatic Cholangiocarcinoma (MSK, Hepatology 2021)

```{r}
Cholangiocarcinoma_patient <- read_csv("Cholangiocarcinoma_data_clinical_patient.csv")
Cholangiocarcinoma_patient <- Cholangiocarcinoma_patient %>% select(PATIENT_ID, AGE, SEX, BMI, OS_MONTHS, OS_STATUS, SMOKING_STATUS) %>% mutate(BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), RACE = NA, TUMOR_PURITY = NA, SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = NA, SITE_OF_TUMOR_TISSUE = "intrahepatic cholangiocarcinoma", METASTATIC_SITE_PATIENT = NA, CANCER_TYPE = "Liver", OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), SMOKING_HISTORY = SMOKING_STATUS, SMOKING_PACK_YEARS = NA, Study = "Intrahepatic Cholangiocarcinoma (MSK, Hepatology 2021)") %>% select(-SMOKING_STATUS)


Cholangiocarcinoma_sample <- read_csv("Cholangiocarcinoma_data_clinical_sample.csv")
Cholangiocarcinoma_sample <- Cholangiocarcinoma_sample %>% select(PATIENT_ID, SAMPLE_ID, TMB_NONSYNONYMOUS) %>% mutate(TMB = TMB_NONSYNONYMOUS) %>% select(-TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n()) # none with more than one sample

Cholangiocarcinoma_data <- merge(Cholangiocarcinoma_patient, Cholangiocarcinoma_sample, by = "PATIENT_ID")
head(Cholangiocarcinoma_data)

# NOTICED LATER: many similarities with previous dataset and other dataset has more info (otherwise same when include info in this study - for about 200 cases)
same_liver_patients <- merge(Hepatocellular_data, Cholangiocarcinoma_data, by = "PATIENT_ID")

Cholangiocarcinoma_data <- Cholangiocarcinoma_data %>% filter(!(PATIENT_ID %in% same_liver_patients$PATIENT_ID))
```

## Kidney Renal Papillary Cell Carcinoma (TCGA, Firehose Legacy)

```{r}
kidney_patient <- read_csv("kidney_data_clinical_patient.csv")
kidney_patient <- kidney_patient %>% select(PATIENT_ID, SEX, RACE, ETHNICITY, AGE, OS_STATUS, OS_MONTHS, SITE_OF_TUMOR_TISSUE, HEIGHT, WEIGHT, SMOKING_PACK_YEARS, TOBACCO_SMOKING_HISTORY_INDICATOR) %>% mutate(BMI = as.numeric(WEIGHT)/(as.numeric(HEIGHT)/100)^2, BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN", "Asian", RACE)), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ETHNICITY), OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), SMOKING_PACK_YEARS = ifelse(SMOKING_PACK_YEARS == "[Not Available]", NA, SMOKING_PACK_YEARS), SMOKING_HISTORY = ifelse(TOBACCO_SMOKING_HISTORY_INDICATOR == "[Not Available]", NA, TOBACCO_SMOKING_HISTORY_INDICATOR), Study = "Kidney Renal Papillary Cell Carcinoma (TCGA, Firehose Legacy)", TUMOR_PURITY = NA, METASTATIC_SITE_PATIENT = NA) %>% select(-c(HEIGHT, WEIGHT, TOBACCO_SMOKING_HISTORY_INDICATOR))

kidney_sample <- read_csv("kidney_data_clinical_sample.csv")
kidney_sample <- kidney_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS) %>% mutate(TMB = TMB_NONSYNONYMOUS, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n()) %>% filter(!(PATIENT_ID == "TCGA-UZ-A9PS" & is.na(TMB) == TRUE)) # one sample same patient but no other TMB/different data

kidney_data <- merge(kidney_patient, kidney_sample, by = "PATIENT_ID")
head(kidney_data)
```

## Skin Cutaneous Melanoma (TCGA, PanCancer Atlas)

```{r}
skin_patient <- read_csv("skin_data_clinical_patient.csv")
skin_patient <- skin_patient %>% select(PATIENT_ID, SEX, HEIGHT, WEIGHT, RACE, ETHNICITY, AGE, METASTATIC_SITE_PATIENT, TUMOR_SITE, OS_STATUS, OS_MONTHS) %>% mutate(BMI = as.numeric(WEIGHT)/(as.numeric(HEIGHT)/100)^2, BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN", "Asian", RACE)), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ETHNICITY), OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), METASTATIC_SITE_PATIENT =  ifelse(METASTATIC_SITE_PATIENT == "[Not Available]", NA, METASTATIC_SITE_PATIENT), SMOKING_HISTORY = NA, SMOKING_PACK_YEARS = NA, SITE_OF_TUMOR_TISSUE = TUMOR_SITE, TUMOR_PURITY = NA, Study = "Skin Cutaneous Melanoma (TCGA, PanCancer Atlas)") %>% select(-c(HEIGHT, WEIGHT, TUMOR_SITE))

skin_sample <- read_csv("skin_data_clinical_sample.csv")
skin_sample <- skin_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS) %>% mutate(TMB = TMB_NONSYNONYMOUS, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n()) %>% filter(!((Code_frequency > 1 & is.na(TMB) == TRUE & is.na(CANCER_TYPE) == TRUE)|(PATIENT_ID == "TCGA-ER-A19T" & is.na(CANCER_TYPE) == TRUE)|(PATIENT_ID == "TCGA-ER-A2NF" & is.na(CANCER_TYPE) == TRUE)))

skin_data <- merge(skin_patient, skin_sample, by = "PATIENT_ID")
head(skin_data)
```

## Uterine Corpus Endometrial Carcinoma (TCGA, Firehose Legacy)

```{r}
uterine_patient <- read_csv("uterine_data_clinical_patient.csv")
uterine_patient <- uterine_patient %>% select(PATIENT_ID, SEX, RACE, ETHNICITY, HEIGHT, WEIGHT, AGE, OS_STATUS, OS_MONTHS, SITE_OF_TUMOR_TISSUE) %>% mutate(BMI = as.numeric(WEIGHT)/(as.numeric(HEIGHT)/100)^2, BMI_CATEGORIES = ifelse(BMI < 18.5, "Underweight", ifelse(BMI < 25, "Normal_Weight", ifelse(BMI < 30, "Overweight", "Obese"))), RACE = ifelse(RACE == "[Not Available]", NA, ifelse(RACE == "ASIAN", "Asian", RACE)), SEX = ifelse(SEX == "[Not Available]", NA, SEX), ETHNICITY = ifelse(ETHNICITY == "[Not Available]", NA, ETHNICITY), OS_STATUS = ifelse(OS_STATUS == "1:DECEASED", 1, 0), METASTATIC_SITE_PATIENT = NA, SMOKING_HISTORY = NA, SMOKING_PACK_YEARS = NA, TUMOR_PURITY = NA, Study = "Uterine Corpus Endometrial Carcinoma (TCGA, Firehose Legacy)") %>% select(-c(HEIGHT, WEIGHT))

uterine_sample <- read_csv("uterine_data_clinical_sample.csv")
uterine_sample <- uterine_sample %>% select(PATIENT_ID, SAMPLE_ID, CANCER_TYPE, TMB_NONSYNONYMOUS) %>% mutate(TMB = TMB_NONSYNONYMOUS, CANCER_TYPE = ifelse(CANCER_TYPE == "[Not Available]", NA, CANCER_TYPE)) %>% select(-TMB_NONSYNONYMOUS) %>% group_by(PATIENT_ID) %>% mutate(Code_frequency= n()) %>% filter(!(PATIENT_ID == "TCGA-BK-A139" & SAMPLE_ID == "TCGA-BK-A139-02")) # same info but no TMB value for other sample

uterine_data <- merge(uterine_patient, uterine_sample, by = "PATIENT_ID")
head(uterine_data)
```

  
# Merge Datasets (all include BMI as well as sex and race for many but smoking needs to have consistent units) 

Issues with smoking: uterine/skin/hepatocellular/esophagogastric no have have, kidney/cervical/bladder has smoking history indicator/smoking history pack years/year started/year stopped, cholangiocarcinoma, Colorectal has smoker status/smoking history indicators, glioblastoma has smoking history/secondhand smoke exposure)
(still working on)

```{r}
BMI_dataset <- rbind(Bladder_data, Esophagogastric_data, Glioblastoma_data, Colorectal_data, Hepatocellular_data, cervical_data, Cholangiocarcinoma_data, kidney_data, skin_data, uterine_data) %>% mutate(SEX = factor(SEX), RACE = factor(RACE), ETHNICITY = factor(ETHNICITY), SITE_OF_TUMOR_TISSUE = factor(SITE_OF_TUMOR_TISSUE), METASTATIC_SITE_PATIENT = factor(METASTATIC_SITE_PATIENT), SMOKING_PACK_YEARS = as.numeric(SMOKING_PACK_YEARS), AGE = as.numeric(AGE), OS_STATUS = factor(OS_STATUS), OS_MONTHS = as.numeric(OS_MONTHS), SMOKING_HISTORY = factor(SMOKING_HISTORY), BMI_CATEGORIES = factor(BMI_CATEGORIES), CANCER_TYPE = factor(CANCER_TYPE), Study = factor(Study), TUMOR_PURITY = as.numeric(TUMOR_PURITY))

# BMI_dataset %>% group_by(PATIENT_ID) %>% mutate(Code_frequency2= n()) %>% filter(Code_frequency2 > 1)
# if same patient but different form of cancer years later how should that be handled 

head(BMI_dataset)

# Current Size: 5989
```

# Summarize/Analyze
```{r}
kable(summary(BMI_dataset))
#BMI_dataset %>% filter(RACE == "BLACK") %>% group_by(Study) %>% summarise(mean(TMB))
```
